\TITLE

### 2020-12-19 response-v3-content.tex

\section*{Reviewer 1}

The revised manuscript is much smoother than the first version, but I
still feel that less than sufficient care was put into the
writing. The reviewer pointed some in the pdf report. I have some
additional editorial suggestions:

\bigskip

  \REVIEW{PROGRESS} {Submit the point-by-point response to the
    reviewer's comments and my comments in a pdf file.}{The original
    submisison web page had a field to enter the comments as text
    which we used. We just added ``see attached PDF'' to make it
    easier for reviewers as suggested by the editor.}

@@

  \REVIEW{DONE} {Needs page number (line number) for ease of
    referencing in the review.}
  {Page numbers and line numbers were
    activated with the review style from the new JDS \LaTeX{}
    template}

  \REVIEW{DONE} {Do not italicize in the text.}
  {All italicized text
    have been removed.  Formulas enclosed in dollar signs are in
    italized text as is default for the JDS document style.}

  \REVIEW{DONE} {AICov needs to be defined in the abstract.}
  {We added ``We present the architecture of an artificial
    intelligence enhanced COVID-19 analysis (in short AICov)''
  in the abstract. We also defined COVID-10 in the abstract.}

\REVIEW{PROGRESS}{Acronyms need to be defined at their first occurrences
  (e.g., AI in the intro). An acronyme table is included in table 1,
  we added the following acronyms to the paper}
  {All acronyms have been defined at their first usage. In addition
    we have added a section Acronyms in which all acronyms are listed
    to make it easier for the reader to look them up in case they are
    used later on in the paper.

### 2020-12-21 vonLaszewski-covid.tex
add local level fixes

\documentclass[letterpaper, inpress]{jds} % use this for production

%\documentclass[a4paper, review]{jds}      % use this for review
\usepackage{lineno}
\linenumbers

\usepackage{calc}
\usepackage{comment}
\usepackage{todonotes}
\setuptodonotes{inline}

@@

\newcommand{\ABSTRACT}{ The COVID-19 (COrona VIrus Disease 2019)
  pandemic has had profound global consequences on health, economic,
  social, behavioral, and almost every major aspect of human
  life. Therefore, it is of great importance to model COVID-19 and
  other pandemics in terms of the broader social contexts in which
  they take place.  We present the architecture of an artificial
  intelligence enhanced COVID-19 analysis (in short AICov) , which
  provides an integrative deep learning framework for COVID-19
  forecasting with population covariates, some of which may serve as
  putative risk factors.  We have integrated multiple different
  strategies into AICov, including the ability to use deep learning
  strategies based on Long Short-Term Memory (LSTM) and event
  modeling.  To demonstrate our approach, we have introduced a
  framework that integrates population covariates from multiple
  sources. Thus, AICov not only includes data on COVID-19 cases and
  deaths but, more importantly, the population's socioeconomic,
  health, and behavioral risk factors at a local level. The compiled
  data are fed into AICov, and thus we obtain improved prediction by
  the integration of the data to our model as compared to one that
  only uses case and death data. As we use deep learning our models
  adapt over time while learning the model from past data.}

@@

For precise and contextually relevant modeling of COVID-19 dynamics
for a given population or community, we use data integration to
combine community-specific health and behavioral risk factors,
demographic and socioeconomic variables that are available at the
local levels along with the cases and deaths data for COVID-19 while
leveraging the recently developed and increasingly popular deep
learning strategies to forecast COVID-19 dynamics using our
model. Towards this, we develop a parallel computing platform called
AICov standing for AI-driven Platform for COVID-19 to implement the
different components of our framework and obtain the needed resources
through multi-cloud interfaces in a robust manner.

@@

In this study, we used a Long Short-Term Memory (LSTM) algorithm for
our predictions \citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a
Recurrent Neural Network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

@@

\begin{requirement}{\bf Interactive Access.}
  Due to the experimental nature of analyzing the data, AICov must
  support the exploration of the data in an interactive
  fashion. \Solution As many data scientists use Jupyter Notebooks to
  integrate their work interactively, AICov integrates such
  notebooks. Jupyter notebooks allow rapid modification of the
  analysis workflow through the ease of Python as programming
  language. Additionally, they enable us to formulate easily
  sophisticated scientific analytics workflows with the help of
  Python.
\end{requirement}

\begin{requirement}{\bf Expandable API.}
  Assuming that new models and other analysis algorithms will be
  developed over time, AICov must allow the integration of such
  APIs. \Solution To address this requirement, we are developed an
  abstraction API for data, analysis, and metadata parameter
  adjustments. External services could be integrated with the help of
  REST (REpresentational State Transfer) services. The REST services
  are automatically generated from function specifications.
\end{requirement}

@@

While the U.S. has seen an overall case fatality rate of 5.14\%, some
U.S. locations have experienced a disproportionate number of deaths
compared to others. This disparity of deaths among different
communities could be attributed to a diversity of local risk factors
and socioeconomic determinants of health that now constitute a very
active area of investigation. Researchers are studying the effects of
different community-specific socioeconomic variables, underlying
health conditions and comorbidities, infrastructure, and systemic
responses.  Given the complexity of such underlying conditions that
can determine individual disease outcomes, where and under what
conditions people get infected by the virus, and then recover or die
take place not in isolation but rather in an interplay of diverse
factors that are characteristic to a local community, which need to be
synthesized. In other words, a basic requirement for effective
modeling of COVID-19 dynamics is that the information regarding a
particular community’s vulnerabilities and strengths pertaining to the
disease must be as comprehensive and integrative as possible.  The
regular reporting of accurate Coronavirus outbreaks data from local
health departments has been understandably difficult, especially in
the areas that are hardest hit. We use data on community-specific
health and behavioral risk factors aggregated at the county level from
the large-scale longitudinal CDC Behavioral Risk Factor Surveillance
System (BRFSS) surveys \citep{www-cdc-brfss}, and demographic and
socioeconomic data from the U.S. Census Bureau that are known at the
local levels along with the cases and deaths data compiled by the CDC
National Center for Health Statistics (NCHS) \citep{www-cdc-nchs} and
integrate them with other data sources.

@@

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 NONE & No risk factor used\\
 ARTHRITIS 	  & Percent reported with arthritis \\
 BINGE 		  & Percent reported with binge drinking \\
 BPHIGH 	  & Percent reported with  high blood pressure\\
 BPMED 		  & Percent reported taking blood pressure medication\\
 CANCER 	  & Percent reported as cancer patients\\
 CASTHMA 	  & Percent reported with current asthma\\
 CHD 		  & Percent reported with coronary heart disease\\
 CHECKUP 	  & Percent reported with health checkup\\
 CHOLSCREEN   & Percent reported with cholesterol screening\\
 COPD 		  & Percent reported with chronic obstructive pulmonary disease\\
 CSMOKING 	  & Percent reported as currently smoking\\
 DIABETES 	  & Percent reported with diabetes\\
 ESTBEDS 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent reported with high cholesterol\\
 INSURANCE 	  & Percent reported with insurance\\
 KIDNEY 	  & Percent reported with kidney disease\\
 LPA 		  & Percent reported with no leisure-time physical activity \\
 MHLTH 		  & Percent reported with not good mental health\\
 NBEDS 		  & Number of beds in hospitals \\
 NBEDS/1000   & Number of beds per 1000 people\\
 NHOSP 		  & Number of hospitals\\
 OBESITY 	  & Percent reported with  population that is obese\\
 PHLTH        & Percent reported with physical health issues\\
 PVI 		  & Pandemic vulnerability index \\
 PERCENTBLACKS 	  & Percent of blacks in the population\\
 PERCENTHISPANICS & Percent of hispanics in the population\\
 STROKE 	  & Percent reported with stroke\\
 BLACK	  & Percent of Blacks in the population\\
 POVERTY & Percent of poverty in the population\\
 SENIOR  & Percent of seniors in the population\\
 SVI\_MINORITY 	  & Percent of social vulnerability index in the minority population\\
 SVI\_OVERALL 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@

To establish the feasibility of using deep learning, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \citep{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered Long Short-Term Memory (LSTM)
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Figure \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\citep{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in both Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

\begin{figure}[!h]
    \centering
    \includegraphics{images/magic-1.png}
    \caption{Cumulative empirical fits can be very accurately
      predicted with LSTM. The data is taken from 37 U.S. Cities. Our
      comparison showcases that given a model, we can replicate the
      prediction of such models using the model data over time. The
      x-axis represents the days since the first data was fed to the
      model.}
    \label{fig:magic-1}
\end{figure}

@@

To answer these questions, we chose for our next analysis a date range
between 2020-02-01 and 2020-05-25. We start with a basic LSTM operator
with two layers of LSTM, and initial and final fully connected layers
\citep{Kadupitiya2020-zq}.  For LSTM, we have selected the activation
function based on Rectified Linear Units (RELU). For the last layer
sigmoid is chosen. A drop out rate of 0.2 is used, we use 16 input
nodes, and have a batch size of 110. The maximum epoch is chosen to be
200. The number of nodes internally is 32. The maximum total number of
Samples is 12540 when using 3 days as input length. We use a two-layer
LSTM, as shown in Table \ref{tab:model}. Within our system, we have
the ability to set a number of input days. For this analysis, we have
chosen three different input lengths, namely 3, 4, and 5 days.

@@

\item[Order for cases sorted by minimum model error:] NONE,
  POP\_DENSITY\_2010, PHLTH, INSURANCE, PERCENTBLACKS,
  PERCENTHISPANICS, NHOSP, INSURANCE, CHOLSCREEN, DIABETES, STROKE,
  CHD, CHECKUP, NBEDS, SVI\_OVERALL, KIDNEY, CASTHMA, BLACK, BPMED,
  LPA, CSMOKING, ARTHRITIS, POVERTY, ESTBEDS, MHLTH, SENIOR, COPD,
  CANCER, NBEDS/1000, BPHIGH, HIGHCHOL, BINGE, OBESITY, SVI\_MINORITY

\item[Order for deaths sorted by minimum model error:] NONE,
  SVI\_OVERALL, CSMOKING, PERCENTHISPANICS, POP\_DENSITY\_2010,
  SENIOR, CHOLSCREEN, CASTHMA, INSURANCE, INSURANCE, MHLTH, LPA,
  POVERTY, OBESITY, CANCER, BLACK, NBEDS/1000, PERCENTBLACKS,
  DIABETES, ARTHRITIS, NBEDS, NHOSP, BPHIGH, CHECKUP, PHLTH, CHD,
  STROKE, HIGHCHOL, ESTBEDS, KIDNEY, BINGE, BPMED, COPD, SVI\_MINORITY
\end{description}

@@

\begin{table}[!p]
\caption{Best models for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.}
\label{tab:top-1}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Cases} &  
 \makecell{RMSE\\Death} &  
 \makecell[r]{Cumu-\\lative \\Error \\Cases} &  
 \makecell[r]{Cumu-\\lative \\Error \\Death} &  
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule
1 & POP\_DENSITY\_2010 & 0.055443 & 0.067617 & 7.94 & 22.74 & 5 \\
2 & PHLTH & 0.054795 & 0.068529 & 8.01 & 20.55 & 3 \\
3 & INSURANCE & 0.055199 & 0.067641 & 8.07 & 20.43 & 4 \\
4 & PERCENTBLACKS & 0.055028 & 0.067207 & 8.12 & 20.17 & 4 \\
5 & PERCENTHISPANICS & 0.056061 & 0.067489 & 8.22 & 19.04 & 5 \\
6 & NHOSP & 0.054621 & 0.068255 & 8.26 & 20.10 & 3 \\
7 & INSURANCE & 0.054557 & 0.068572 & 8.34 & 24.61 & 3 \\
8 & CHOLSCREEN & 0.054712 & 0.069121 & 8.37 & 21.70 & 3 \\
9 & DIABETES & 0.054586 & 0.067636 & 8.40 & 20.50 & 5 \\
10 & STROKE & 0.054160 & 0.066906 & 8.45 & 20.16 & 4 \\
11 & CHD & 0.054027 & 0.067642 & 8.46 & 21.93 & 5 \\
12 & CHECKUP & 0.054927 & 0.067652 & 8.51 & 20.42 & 5 \\
13 & NBEDS & 0.054146 & 0.067536 & 8.54 & 21.15 & 5 \\
14 & SVI\_OVERALL & 0.054795 & 0.067676 & 8.58 & 19.94 & 5 \\
15 & KIDNEY & 0.054600 & 0.067565 & 8.61 & 21.86 & 5 \\
16 & CASTHMA & 0.055089 & 0.068248 & 8.63 & 22.58 & 5 \\
17 & BLACK & 0.054547 & 0.068266 & 8.68 & 20.49 & 3 \\
18 & BPMED & 0.054685 & 0.068920 & 8.74 & 21.69 & 3 \\
19 & LPA & 0.054323 & 0.067149 & 8.75 & 19.54 & 5 \\
20 & CSMOKING & 0.053716 & 0.066898 & 8.75 & 19.01 & 5 \\
21 & ARTHRITIS & 0.054053 & 0.068716 & 8.83 & 20.94 & 3 \\
22 & POVERTY & 0.055029 & 0.068672 & 8.84 & 22.02 & 3 \\
23 & ESTBEDS & 0.054545 & 0.067267 & 8.86 & 21.22 & 4 \\
24 & MHLTH & 0.054864 & 0.068742 & 8.90 & 19.53 & 3 \\
25 & SENIOR & 0.055079 & 0.069024 & 8.93 & 21.23 & 3 \\
26 & COPD & 0.055589 & 0.068250 & 8.93 & 23.26 & 4 \\
27 & CANCER & 0.054500 & 0.068363 & 8.99 & 19.70 & 3 \\
28 & NBEDS/1000 & 0.054397 & 0.068491 & 9.05 & 21.89 & 3 \\
29 & BPHIGH & 0.054609 & 0.067346 & 9.13 & 21.29 & 4 \\
\rowcolor{black!5} 30 & NONE & 0.054868 & 0.069432 & 9.26 & 25.88 & 3 \\
31 & HIGHCHOL & 0.056429 & 0.069345 & 9.28 & 20.98 & 3 \\
32 & BINGE & 0.054317 & 0.067638 & 9.32 & 21.00 & 5 \\
33 & OBESITY & 0.054469 & 0.068540 & 9.50 & 20.65 & 3 \\
34 & SVI\_MINORITY & 0.055094 & 0.067443 & 9.54 & 21.53 & 4 \\
\bottomrule
\end{tabular}
%}
\end{table}

@@

HISPANICS               & -0.072      & -0.032      & -0.064      & -0.091      & -0.068      & -0.046    \\

>>>>

\documentclass[letterpaper, inpress]{jds} % use this for production
%\documentclass[a4paper, review]{jds}      % use this for review
\usepackage{lineno}
\linenumbers

\usepackage{subcaption}
\usepackage{calc}
\usepackage{comment}
\usepackage{todonotes}
\setuptodonotes{inline}

@@

\newcommand{\ABSTRACT}{ The COVID-19 (COrona VIrus Disease 2019)
  pandemic has had profound global consequences on health, economic,
  social, behavioral, and almost every major aspect of human
  life. Therefore, it is of great importance to model COVID-19 and
  other pandemics in terms of the broader social contexts in which
  they take place.  We present the architecture of an artificial
  intelligence enhanced COVID-19 analysis (in short AICov) , which
  provides an integrative deep learning framework for COVID-19
  forecasting with population covariates, some of which may serve as
  putative risk factors.  We have integrated multiple different
  strategies into AICov, including the ability to use deep learning
  strategies based on Long Short-Term Memory (LSTM) and event
  modeling.  To demonstrate our approach, we have introduced a
  framework that integrates population covariates from multiple
  sources. Thus, AICov not only includes data on COVID-19 cases and
  deaths but, more importantly, the population's socioeconomic,
  health, and behavioral risk factors at their specific locations.
  The compiled data are fed into AICov, and thus we obtain improved
  prediction by the integration of the data to our model as compared
  to one that only uses case and death data. As we use deep learning
  our models adapt over time while learning the model from past data.}

@@

For precise and contextually relevant modeling of COVID-19 dynamics
for a given population or community, we use data integration to
combine community-specific health and behavioral risk factors,
demographic and socioeconomic variables that are available at the
local along with the cases and deaths data for COVID-19 while
leveraging the recently developed and increasingly popular deep
learning strategies to forecast COVID-19 dynamics using our
model. Towards this, we develop a parallel computing platform called
AICov standing for AI-driven Platform for COVID-19 to implement the
different components of our framework and obtain the needed resources
through multi-cloud interfaces in a robust manner.

@@

In this study, we used a LSTM algorithm for
our predictions \citep{www-keras-lstm,Hochreiter1997-dk}. An LSTM is a
Recurrent Neural Network (RNN) \citep{Rumelhart1986-li} with feedback
connections allowing the use of data input sequences to predict data
output sequences. LSTMs have been applied to many application areas
from handwriting \citep{Graves2009-qb}, image, feature detection, and
time series prediction \citep{Schmidhuber2005-oy}. LSTM's have an
internal state that is used to prevent the vanishing gradient problem
\citep{Hochreiter1991-mp} during the training of RNNs.

@@

\begin{requirement}{\bf Interactive Access.}
  Due to the experimental nature of analyzing the data, AICov must
  support the exploration of the data in an interactive
  fashion. \Solution As many data scientists use Jupyter Notebooks to
  integrate their work interactively, AICov integrates such
  notebooks. Jupyter notebooks allow rapid modification of the
  analysis workflow through the ease of Python as programming
  language. Additionally, they enable us to formulate easily
  sophisticated scientific analytics workflows with the help of
  Python and Jupyter Notebooks.
\end{requirement}

\begin{requirement}{\bf Expandable API.}
  Assuming that new models and other analysis algorithms will be
  developed over time, AICov must allow the integration of such
  APIs. \Solution To address this requirement, we are developing an
  abstract API for data, analysis, and metadata parameter
  adjustments. External services can be integrated with the help of
  REST (REpresentational State Transfer) services. The REST services
  are automatically generated from function specifications.
\end{requirement}

@@

While the U.S. has seen an overall case fatality rate of 5.14\%, some
U.S. locations have experienced a disproportionate number of deaths
compared to others. This disparity of deaths among different
communities could be attributed to a diversity of local risk factors
and socioeconomic determinants of health that now constitute a very
active area of investigation. Researchers are studying the effects of
different community-specific socioeconomic variables, underlying
health conditions and comorbidities, infrastructure, and systemic
responses.  Given the complexity of such underlying conditions that
can determine individual disease outcomes, where and under what
conditions people get infected by the virus, and then recover or die
take place not in isolation but rather in an interplay of diverse
factors that are characteristic to a local community, which need to be
synthesized. In other words, a basic requirement for effective
modeling of COVID-19 dynamics is that the information regarding a
particular community’s vulnerabilities and strengths pertaining to the
disease must be as comprehensive and integrative as possible.  The
regular reporting of accurate Coronavirus outbreaks data from local
health departments has been understandably difficult, especially in
the areas that are hardest hit.
%
\linelabel{comment:level} We use data on community-specific health and
behavioral risk factors aggregated at the county level from the
large-scale longitudinal CDC Behavioral Risk Factor Surveillance
System (BRFSS) surveys \citep{www-cdc-brfss} and demographic and
socioeconomic data from the U.S. Census Bureau \cite{www-sensus},
i.e., both at the county and census tract, along with the cases and
deaths data compiled by the CDC National Center for Health Statistics
(NCHS) \citep{www-cdc-nchs} and integrate them with other data
sources. These other data sources include

\todo{Pyne}.

@@

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 NONE & No risk factor used\\
 ARTHRITIS 	  & Percent reported with arthritis \\
 BINGE 		  & Percent reported with binge drinking \\
 BPHIGH 	  & Percent reported with  high blood pressure\\
 BPMED 		  & Percent reported taking blood pressure medication\\
 CANCER 	  & Percent reported as cancer patients\\
 CASTHMA 	  & Percent reported with current asthma\\
 CHD 		  & Percent reported with coronary heart disease\\
 CHECKUP 	  & Percent reported with health checkup\\
 CHOLSCREEN   & Percent reported with cholesterol screening\\
 COPD 		  & Percent reported with chronic obstructive pulmonary disease\\
 CSMOKING 	  & Percent reported as currently smoking\\
 DIABETES 	  & Percent reported with diabetes\\
 ESTBEDS 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent reported with high cholesterol\\
 INSURANCE 	  & Percent reported with insurance\\
 KIDNEY 	  & Percent reported with kidney disease\\
 LPA 		  & Percent reported with no leisure-time physical activity \\
 MHLTH 		  & Percent reported with not good mental health\\
 NBEDS 		  & Number of beds in hospitals \\
 NBEDS/1000   & Number of beds per 1000 people\\
 NHOSP 		  & Number of hospitals\\
 OBESITY 	  & Percent reported with  population that is obese\\
 PHLTH        & Percent reported with physical health issues\\
 PVI 		  & Pandemic vulnerability index \\
 STROKE 	  & Percent reported with stroke\\
 BLACK	  & Percent of Blacks in the population\\
 POVERTY & Percent of poverty in the population\\
 SENIOR  & Percent of seniors in the population\\
 SVI\_MINORITY 	  & Percent of social vulnerability index in the minority population\\
 SVI\_OVERALL 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@

To establish the feasibility of using deep learning, we have, in our
analysis, verified that through deep learning we can recreate the
predictions identified by a model-based approach using empirical
fits. In this comprehensive analysis, we obtain predictions through
deep learning while using data generated from epidemiological models
as discussed in \citep{marsland20-covid-paper}. The time series we used
was 100 days long, and a multi-layered LSTM
recurrent network was used. Our prediction approach differs by
learning not only from the demographics (fixed data for each city) and
time-dependent data but by integrating the population model for the
underlying prediction, as shown in Figure \ref{fig:magic-1}. Such model
predictive integration capability is important in any application with
multiple time scales. For example, this allows us to integrate
multi-scale time effects into the forecast, which could address the
combination of general forecasts and the next time steps of choice,
such as days, weeks, months, or longer. For this analysis, we used 37
of the 110 cities with reliable empirical fits (that we derived from
an epidemiological model and not deep learning
\citep{marsland20-covid-paper}) for the cases and deaths data up to
April 15, 2020. Our analysis identified a sophisticated single deep
learning time evolution operator that can describe these 37 separate
data sets, and smooth fitted data leads to very accurate deep learning
descriptions via LSTM as depicted in both Figure \ref{fig:magic-1}
that are smaller than 1\% from the model predictions.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{images/magic-1.png}

    \begin{subfigure}{.4\textwidth}
      \centering
      % include first image
      % \includegraphics[width=.8\linewidth]{image_file_name}  
      \caption{Prediction acuracy for deaths.}
      \label{fig:magic-a}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
      \centering
      % include second image
      %\includegraphics[width=.8\linewidth]{image_file_name}  
      \caption{Prediction acuracy for cases.}
      \label{fig:magic-b}
    \end{subfigure}

    \caption{Cumulative empirical fits can be very accurately
      predicted with LSTM. The data is taken from 37 U.S. Cities. Our
      comparison showcases that given a model, we can replicate the
      prediction of such models using the model data over time. The
      x-axis represents the days since the first data was fed to the
      model.}
    \label{fig:magic-1}
\end{figure}

@@

To answer these questions, we chose for our next analysis a date range
between 2020-02-01 and 2020-05-25. We start with a basic LSTM operator
with two layers of LSTM, and initial and final fully connected layers
\citep{Kadupitiya2020-zq}.  For LSTM, we have selected the activation
function based on Rectified Linear Units (RELU). For the last layer
sigmoid is chosen. A drop out rate of 0.2 is used, we use 16 input
nodes, and have a batch size of 110. The maximum epoch is chosen to be
200. The number of nodes internally is 32. The maximum total number of
samples is 12540 when using 3 days as input length. We use a two-layer
LSTM, as shown in Table \ref{tab:model}. Within our system, we have
the ability to set a number of input days. For this analysis, we have
chosen three different input lengths, namely 3, 4, and 5 days. The
term ``None'' in the output shape means this dimension is
variable. For our model it is the batch size.  \todo{What is the batch
size.}

@@

\item[Order for cases sorted by minimum model error:] NONE,
  POP\_DENSITY\_2010, PHLTH, INSURANCE, BLACK,
  HISPANICS, NHOSP, CHOLSCREEN, DIABETES, STROKE,
  CHD, CHECKUP, NBEDS, SVI\_OVERALL, KIDNEY, CASTHMA, BPMED,
  LPA, CSMOKING, ARTHRITIS, POVERTY, ESTBEDS, MHLTH, SENIOR, COPD,
  CANCER, NBEDS/1000, BPHIGH, HIGHCHOL, BINGE, OBESITY, SVI\_MINORITY

\item[Order for deaths sorted by minimum model error:] NONE,
  SVI\_OVERALL, CSMOKING, HISPANICS, POP\_DENSITY\_2010,
  SENIOR, CHOLSCREEN, CASTHMA, INSURANCE, MHLTH, LPA,
  POVERTY, OBESITY, CANCER, BLACK, NBEDS/1000, DIABETES, ARTHRITIS,
  NBEDS, NHOSP, BPHIGH, CHECKUP, PHLTH, CHD, STROKE, HIGHCHOL,
  ESTBEDS, KIDNEY, BINGE, BPMED, COPD, SVI\_MINORITY
\end{description}

@@

\begin{table}[!p]
\caption{Best models for each risk factor and their respective
  errors. The model that uses no risk factor is highlighted in grey.}
\label{tab:top-1}
\bigskip
\centering
%\resizebox{1.0\textwidth}{!}{ 
\begin{tabular}{rrrrrlrl}
\toprule
 \makecell{Place/\\Rank} & 
\makecell{Risk Factor} &  
\makecell{RMSE\\Cases} &  
 \makecell{RMSE\\Death} &  
 \makecell[r]{Cumu-\\lative \\Error \\Cases} &  
 \makecell[r]{Cumu-\\lative \\Error \\Death} &  
 \makecell[r]{Number of \\Days as \\Input} \\
\midrule
1 & POP\_DENSITY\_2010 & 0.055443 & 0.067617 & 7.94 & 22.74 & 5 \\
2 & PHLTH & 0.054795 & 0.068529 & 8.01 & 20.55 & 3 \\
3 & INSURANCE & 0.055199 & 0.067641 & 8.07 & 20.43 & 4 \\
4 & BLACK & 0.055028 & 0.067207 & 8.12 & 20.17 & 4 \\
5 & HISPANIC & 0.056061 & 0.067489 & 8.22 & 19.04 & 5 \\
6 & NHOSP & 0.054621 & 0.068255 & 8.26 & 20.10 & 3 \\
7 & CHOLSCREEN & 0.054712 & 0.069121 & 8.37 & 21.70 & 3 \\
8 & DIABETES & 0.054586 & 0.067636 & 8.40 & 20.50 & 5 \\
9 & STROKE & 0.054160 & 0.066906 & 8.45 & 20.16 & 4 \\
10 & CHD & 0.054027 & 0.067642 & 8.46 & 21.93 & 5 \\
11 & CHECKUP & 0.054927 & 0.067652 & 8.51 & 20.42 & 5 \\
12 & NBEDS & 0.054146 & 0.067536 & 8.54 & 21.15 & 5 \\
13 & SVI\_OVERALL & 0.054795 & 0.067676 & 8.58 & 19.94 & 5 \\
14 & KIDNEY & 0.054600 & 0.067565 & 8.61 & 21.86 & 5 \\
15 & CASTHMA & 0.055089 & 0.068248 & 8.63 & 22.58 & 5 \\
16 & BPMED & 0.054685 & 0.068920 & 8.74 & 21.69 & 3 \\
17 & LPA & 0.054323 & 0.067149 & 8.75 & 19.54 & 5 \\
18 & CSMOKING & 0.053716 & 0.066898 & 8.75 & 19.01 & 5 \\
19 & ARTHRITIS & 0.054053 & 0.068716 & 8.83 & 20.94 & 3 \\
20 & POVERTY & 0.055029 & 0.068672 & 8.84 & 22.02 & 3 \\
21 & ESTBEDS & 0.054545 & 0.067267 & 8.86 & 21.22 & 4 \\
22 & MHLTH & 0.054864 & 0.068742 & 8.90 & 19.53 & 3 \\
23 & SENIOR & 0.055079 & 0.069024 & 8.93 & 21.23 & 3 \\
24 & COPD & 0.055589 & 0.068250 & 8.93 & 23.26 & 4 \\
25 & CANCER & 0.054500 & 0.068363 & 8.99 & 19.70 & 3 \\
26 & NBEDS/1000 & 0.054397 & 0.068491 & 9.05 & 21.89 & 3 \\
27 & BPHIGH & 0.054609 & 0.067346 & 9.13 & 21.29 & 4 \\
\rowcolor{black!5} 28 & NONE & 0.054868 & 0.069432 & 9.26 & 25.88 & 3 \\
29 & HIGHCHOL & 0.056429 & 0.069345 & 9.28 & 20.98 & 3 \\
30 & BINGE & 0.054317 & 0.067638 & 9.32 & 21.00 & 5 \\
31 & OBESITY & 0.054469 & 0.068540 & 9.50 & 20.65 & 3 \\
32 & SVI\_MINORITY & 0.055094 & 0.067443 & 9.54 & 21.53 & 4 \\
\bottomrule
\end{tabular}
%}
\end{table}

@@

HISPANIC                & -0.072      & -0.032      & -0.064      & -0.091      & -0.068      & -0.046    \\
